Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)

### 2025-04-28 02-antecedents.qmd
Update results, method and render AL

H1: Lower perceived economic inequality is associated with stronger market justice preferences.

H2: Stronger meritocratic beliefs are associated with stronger market justice preferences.

H3: The positive association between meritocratic beliefs and market justice preferences is moderated by perceived economic inequality; specifically, this association is weaker when perceived economic inequality is high.

H4: The effects described in H1–H3 are attenuated after major social mobilizations (2019–2022), reflecting increased critical views of the market's role in allocating social goods.

>>>>

$H1$: Greater perceived economic inequality is associated with stronger market justice preferences.

$H2$: Stronger meritocratic beliefs are associated with stronger market justice preferences.

$H3$: The positive association between meritocratic beliefs and market justice preferences is moderated by perceived economic inequality; specifically, this association is weaker when perceived economic inequality is high.

$H4$: The effects described in H1–H3 are attenuated after major social mobilizations (2019–2022), reflecting increased critical views of the market's role in allocating social goods.

### 2025-04-28 03-methods.qmd
Update results, method and render AL

df_study1 <- df_study1_long_t7 %>%
  select(idencuesta,
         ola,
         ponderador_long_total,
         segmento,
         estrato,
         just_health,
         just_pension,
         just_educ,
         mjp, 
         perc_inequality, 
         just_inequality,
         merit_effort, 
         merit_talent, 
         educ,
         quintil1,
         sex,
         age,
         ess, 
         ideo) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

@@

**Perception of economic inequality**: The main independent variable refers to the perception of economic inequality, measured through the perceived wage gap. [@castillo_perception_2022] [acá citar a Hadler, Gijsberts, Castillo 2009 (CEP)]. This measure is derived from the salary gap between the perceived salaries of jobs at opposite ends of the occupational hierarchy. Specifically, it relies on the division between the perceived salary of a large-company president and that of an unskilled worker [@Castillo2011]. Higher values of this term indicate a greater perception of economic inequality between occupations located at the extremes of the status continuum. This measure includes a logarithmic term in order to adjust income magnitudes (usually fewer cases with high income):

@@

t2 <- df_study1 %>% 
  filter(ola == 1) %>% 
  select(perc_inequality, just_inequality, merit_effort, merit_talent)

>>>>

df_study1 <- df_study1_long_t7 %>%
  select(idencuesta,
         ola,
         ponderador_long_total,
         segmento,
         estrato,
         just_health,
         just_pension,
         just_educ,
         mjp, 
         perc_inequality, 
         merit_effort, 
         merit_talent, 
         educ,
         quintil1,
         sex,
         age,
         ess, 
         ideo) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

@@

**Perception of economic inequality**: The main independent variable refers to the perception of economic inequality, measured through the perceived wage gap. [@castillo_cual_2009; @hadler_why_2005; @gijsberts_thelegitimation_1999]. This measure is derived from the salary gap between the perceived salaries of jobs at opposite ends of the occupational hierarchy. Specifically, it relies on the division between the perceived salary of a large-company president and that of an unskilled worker [@Castillo2011]. Higher values of this term indicate a greater perception of economic inequality between occupations located at the extremes of the status continuum. This measure includes a logarithmic term in order to adjust income magnitudes (usually fewer cases with high income):

@@

t2 <- df_study1 %>% 
  filter(ola == 1) %>% 
  select(perc_inequality, merit_effort, merit_talent)
